---
title: "Master Dataset"
author: "Yuki Joyama"
date: "2023-11-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Use codes in this document to import the master dataset.

- NYPD Complaint Data Historic (NYC OpenData)\
- Data collected between 2017 Jan 01 12:00:00 AM - 2022 Dec 31 11:45:00 PM\
[data source](https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Data-Historic/qgea-i56i/explore/query/SELECT%0A%20%20%60cmplnt_num%60%2C%0A%20%20%60cmplnt_fr_dt%60%2C%0A%20%20%60cmplnt_fr_tm%60%2C%0A%20%20%60cmplnt_to_dt%60%2C%0A%20%20%60cmplnt_to_tm%60%2C%0A%20%20%60addr_pct_cd%60%2C%0A%20%20%60rpt_dt%60%2C%0A%20%20%60ky_cd%60%2C%0A%20%20%60ofns_desc%60%2C%0A%20%20%60pd_cd%60%2C%0A%20%20%60pd_desc%60%2C%0A%20%20%60crm_atpt_cptd_cd%60%2C%0A%20%20%60law_cat_cd%60%2C%0A%20%20%60boro_nm%60%2C%0A%20%20%60loc_of_occur_desc%60%2C%0A%20%20%60prem_typ_desc%60%2C%0A%20%20%60juris_desc%60%2C%0A%20%20%60jurisdiction_code%60%2C%0A%20%20%60parks_nm%60%2C%0A%20%20%60hadevelopt%60%2C%0A%20%20%60housing_psa%60%2C%0A%20%20%60x_coord_cd%60%2C%0A%20%20%60y_coord_cd%60%2C%0A%20%20%60susp_age_group%60%2C%0A%20%20%60susp_race%60%2C%0A%20%20%60susp_sex%60%2C%0A%20%20%60transit_district%60%2C%0A%20%20%60latitude%60%2C%0A%20%20%60longitude%60%2C%0A%20%20%60lat_lon%60%2C%0A%20%20%60patrol_boro%60%2C%0A%20%20%60station_name%60%2C%0A%20%20%60vic_age_group%60%2C%0A%20%20%60vic_race%60%2C%0A%20%20%60vic_sex%60%2C%0A%20%20%60%3A%40computed_region_efsh_h5xi%60%2C%0A%20%20%60%3A%40computed_region_f5dn_yrer%60%2C%0A%20%20%60%3A%40computed_region_yeji_bk3q%60%2C%0A%20%20%60%3A%40computed_region_92fq_4b7q%60%2C%0A%20%20%60%3A%40computed_region_sbqj_enih%60%0AWHERE%0A%20%20%60cmplnt_fr_dt%60%0A%20%20%20%20BETWEEN%20%222017-01-01T00%3A00%3A00%22%20%3A%3A%20floating_timestamp%0A%20%20%20%20AND%20%222022-12-31T23%3A45%3A00%22%20%3A%3A%20floating_timestamp%0AORDER%20BY%20%60rpt_dt%60%20DESC%20NULL%20FIRST/page/filter)  

## Import NYPD csv file from dropbox

```{r}
df_nypd = read_csv("https://www.dropbox.com/scl/fi/kf2zk4t1onxzm2vo3lpkq/NYPD_Complaint_Data_Historic.csv?rlkey=ly36vi9v66sno80eir6rohlwn&dl=1", na = "(null)") |> # some values are coded as "(null)" in the df; rewrite them as NA
  janitor::clean_names() 
```

The file size is 0.97GB. The dataset should have `r ncol(df_nypd)` variables and `r nrow(df_nypd)` rows.\

Column names are changed to lower cases using `janitor::clean_names()`. Please refer to this [link](https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Data-Historic/qgea-i56i) for more details about columns. 

## Socioeconomic indicators 

```{r}
df_socio = readxl::read_excel(here::here("data", "neighorhood_Indicators.xlsx"), sheet = "Data")
```

This dataset (`r ncol(df_socio)` variables and `r nrow(df_socio)` rows) provides information on the NYC socioecomonic background. ([data source](https://furmancenter.org/coredata/userguide/about))\
The excel file is stored in "data" directory. Variable explanations can be found in the "Data Dictionary" sheet.

## Neighborhoods (Sub-borough) Information

The neighborhoods information is not included in our master dataset so we will add them using `addr_pct_cd` (NYPD precinct codes) and approximate mapping from Golub et al. (2006)  
NYPD the 121st Precinct serves the northwestern shore of Staten Island (Willowbrook, Westerleigh, Port Richmond, Mariner's Harbor, Elm Park, Port Ivory, Chelsea, and Bloomfield) so we classified this as `North Shore`.

```{r}
df_nypd_loc = df_nypd |> 
  mutate(
    region_name = case_when( # region_name: neighborhood, sub-borough
      # manhattan
      addr_pct_cd %in% c(1, 6) ~ "Greenwich Village/Financial District", 
      addr_pct_cd %in% c(5, 7, 9)  ~ "Lower East Side/Chinatown",
      addr_pct_cd %in% c(10, 18, 14)  ~ "Chelsea/Clinton/Midtown",
      addr_pct_cd %in% c(13, 17) ~ "Stuyvesant Town/Turtle Bay",
      addr_pct_cd %in% c(20, 24) ~ "Upper West Side",
      addr_pct_cd == 19 ~ "Upper East Side",
      addr_pct_cd == 22 ~ "Central Park",
      addr_pct_cd %in% c(26, 30) ~ "Morningside Heights/Hamilton Heights",
      addr_pct_cd %in% c(28, 32) ~ "Central Harlem",
      addr_pct_cd %in% c(23, 25) ~ "East Harlem",
      addr_pct_cd %in% c(33, 34) ~ "Washington Heights/Inwood",
      # the bronx
      addr_pct_cd %in% c(40, 41) ~ "Mott Haven/Hunts Point",
      addr_pct_cd %in% c(42, 48) ~ "Morrisania/Belmont",
      addr_pct_cd == 44 ~ "Highbridge/South Concourse",
      addr_pct_cd == 46 ~ "University Heights/Fordham",
      addr_pct_cd == 52 ~ "Kingsbridge Heights/Mosholu", 
      addr_pct_cd == 50 ~ "Riverdale/Kingsbridge" ,
      addr_pct_cd == 43 ~ "Soundview/Parkchester",
      addr_pct_cd == 45 ~ "Throgs Neck/Co-op City",
      addr_pct_cd == 49 ~ "Pelham Parkway",
      addr_pct_cd == 47 ~ "Williamsbridge/Baychester",
      # brooklyn
      addr_pct_cd %in% c(90, 94) ~ "Williamsburg/Greenpoint",
      addr_pct_cd %in% c(84, 88) ~ "Brooklyn Heights/Fort Greene",
      addr_pct_cd %in% c(79, 81) ~ "Bedford Stuyvesant",
      addr_pct_cd == 83 ~ "Bushwick",
      addr_pct_cd == 75 ~ "East New York/Starrett City",
      addr_pct_cd %in% c(76, 78) ~ "Park Slope/Carroll Gardens",
      addr_pct_cd == 72 ~ "Sunset Park",
      addr_pct_cd == 77 ~ "North Crown Heights/Prospect Heights",
      addr_pct_cd == 71 ~ "South Crown Heights",
      addr_pct_cd == 68 ~ "Bay Ridge",
      addr_pct_cd == 62 ~ "Bensonhurst" ,
      addr_pct_cd == 66 ~ "Borough Park",
      addr_pct_cd == 60 ~ "Coney Island",
      addr_pct_cd == 70 ~ "Flatbush",
      addr_pct_cd == 61 ~ "Sheepshead Bay/Gravesend",
      addr_pct_cd == 73 ~ "Brownsville/Ocean Hill",
      addr_pct_cd == 67 ~ "East Flatbush",
      addr_pct_cd %in% c(63, 69) ~ "Flatlands/Canarsie",
      # queens
      addr_pct_cd == 114 ~ "Astoria",
      addr_pct_cd == 108 ~ "Sunnyside/Woodside",
      addr_pct_cd == 115 ~ "Jackson Heights",
      addr_pct_cd == 110 ~ "Elmhurst/Corona",
      addr_pct_cd == 104 ~ "Middle Village/Ridgewood",
      addr_pct_cd == 112 ~ "Rego Park/Forest Hills",
      addr_pct_cd == 109 ~ "Flushing/Whitestone",
      addr_pct_cd == 107 ~ "Hillcrest/Fresh Meadows",
      addr_pct_cd == 102 ~ "Ozone Park/Woodhaven",
      addr_pct_cd == 106 ~ "South Ozone Park/Howard Beach", 
      addr_pct_cd == 111 ~ "Bayside/Little Neck",
      addr_pct_cd %in% c(103, 113) ~ "Jamaica",
      addr_pct_cd == 105 ~ "Queens Village",
      addr_pct_cd %in% c(100, 101) ~ "Rockaways",
      # staten island
      addr_pct_cd %in% c(120, 121) ~ "North Shore",
      addr_pct_cd == 122 ~ "Mid-Island",
      addr_pct_cd == 123 ~ "South Shore"
    )
  ) |> 
 drop_na(region_name) # excluded the data without the information of location 
```
