---
title: "Siqing analysis"
author: "Siqing Wang"
date: "2023-12-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message=FALSE,
  warning=FALSE)
library(tidyverse)
library(tidyr)
library(dplyr)
```

```{r,include=FALSE}
# df_nypd = read_csv("https://www.dropbox.com/scl/fi/kf2zk4t1onxzm2vo3lpkq/NYPD_Complaint_Data_Historic.csv?rlkey=ly36vi9v66sno80eir6rohlwn&dl=1", na = "(null)") |> # some values are coded as "(null)" in the df; rewrite them as NA
#   janitor::clean_names() 
df_nypd = read_csv("../NYPD_Complaint_Data_Historic.csv") |> 
  janitor::clean_names()
```

Reading in the dataset that map each precinct to a neighborhood, for further SES analysis
```{r}
prec_neighbor = read_csv("data/nyc_prec_neighborhood.csv")
```

Reading in the SES data for demographics information for each precinct/neighboorhood, and only keeping the relevant SES indicators 

```{r}
neighbor_ses = readxl::read_excel("data/neighorhood_indicators.xlsx", sheet = "Data") |> 
  janitor::clean_names() |> 
  filter(region_type == "Sub-Borough Area") |> 
  rename(neighborhood = region_name) |> 
  select(neighborhood, year, pop_num, gross_rent_0_1beds, gross_rent_2_3beds, hh_inc_med_adj, pop16_unemp_pct, pop_edu_collp_pct, pop_edu_nohs_pct, pop_pov_pct, pop_race_asian_pct, pop_race_black_pct, pop_race_hisp_pct, pop_race_white_pct, pop_foreign_pct) |> 
  filter(year %in% c(2017, 2018, 2019, 2020, 2021, 2022))
```

`neighbor_ses` df includes information on population, income, umemployment, poverty, education, and racial factors for every neighborhood. 

Merging `prec_neighboor` with NYPD dataset, conducting more data cleaning. The resulting `nypd_ses_df` only contains variables of interest for SES analysis. 
```{r}
nypd_ses_df = df_nypd |> 
  select(cmplnt_num, cmplnt_fr_dt, addr_pct_cd, crm_atpt_cptd_cd, law_cat_cd, boro_nm, susp_age_group, susp_race, susp_sex, vic_age_group, vic_race, vic_sex, ofns_desc, pd_desc) |> 
  rename(precinct = addr_pct_cd) |> 
  mutate(cmplnt_fr_dt = as.Date(cmplnt_fr_dt, format = "%m/%d/%Y"),
         year = format(cmplnt_fr_dt, "%Y")) |> 
  left_join(prec_neighbor, by = "precinct")
```

`nypd_ses_df` now contains neighboorhood information, as well as other necessary data from the original master data 

Now, calculate the crime rate for every precinct. Crime rate is calculated by dividing the number of crimes committed per neighborhood over the total population of that neighborhood, then times 100,000. Then merge the crime rate data with other relevant socioeconomic data. We focus only on the 2021 data. 
```{r}
ses_df = nypd_ses_df |> 
  group_by(year, neighborhood) |> 
  summarise(crime_num = n())

ses_df = ses_df |> merge(neighbor_ses, by = c("year", "neighborhood")) |> 
  mutate(crime_rate = (crime_num/pop_num) * 100,000)
```
`ses_df` contains crime rate and other SES data per neighborhood, by year 

Visualizing trend: are crime rates associated with poverty, income, and educational level? 

```{r}
eda_plot1 = ses_df |> ggplot(aes(x = pop_pov_pct, y = crime_rate)) +
  geom_point() +  labs(title = "Population in Poverty vs. Crime Rate")

eda_plot2 = ses_df |> ggplot(aes(x = hh_inc_med_adj, y = crime_rate)) +
  geom_point() + labs(title = "Median Household Income vs. Crime Rate")

eda_plot3 = ses_df |> ggplot(aes(x = pop16_unemp_pct, y = crime_rate)) +
  geom_point() + labs(title = "Population Unemployment Percentage vs. Crime Rate")

eda_plot4 = ses_df |> ggplot(aes(x = pop_edu_nohs_pct, y = crime_rate)) +
  geom_point() + labs(title = "Population with No High School Education vs. Crime Rate")

gridExtra::grid.arrange(
  eda_plot1, eda_plot2, eda_plot3, eda_plot4, ncol = 2)
```

Visualizing trend: are crime rates associated with racial factors?
```{r}
eda_plot5 = ses_df |> ggplot(aes(x = pop_race_asian_pct, y = crime_rate)) +
  geom_point() +  labs(title = "Asian% in Poverty vs. Crime Rate")

eda_plot6 = ses_df |> ggplot(aes(x = pop_race_black_pct, y = crime_rate)) +
  geom_point() +  labs(title = "Black% in Poverty vs. Crime Rate")

eda_plot7 = ses_df |> ggplot(aes(x = pop_race_hisp_pct, y = crime_rate)) +
  geom_point() +  labs(title = "Hispanic% in Poverty vs. Crime Rate")

eda_plot8 = ses_df |> ggplot(aes(x = pop_race_white_pct, y = crime_rate)) +
  geom_point() +  labs(title = "White% in Poverty vs. Crime Rate")

gridExtra::grid.arrange(
  eda_plot5, eda_plot6, eda_plot7, eda_plot8, ncol = 2)
```





