---
title: "SES Analysis"
output: html_document
---

We are interested in exploring whether socioeconomic factors are associated or can be used as predictors to assess crime rate. Since our NYPD complaint dataset only contains individual complaints, we utilized a socioeconomic indicator dataset (INSERT LINK: NYU Furman Center) to gather population data for each borough and neighborhood. 

To better conduct regression analysis, we have calculated the crime rate for each neighborhood as the following:
Crime Rate = (Total number of complaints in a neighborhood in a given year / Total population in a neighborhood in a given year) * 100,000

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message=FALSE,
  warning=FALSE)
library(tidyverse)
library(tidyr)
library(dplyr)
library(glmnet)
```

```{r}
# df_nypd = read_csv("https://www.dropbox.com/scl/fi/kf2zk4t1onxzm2vo3lpkq/NYPD_Complaint_Data_Historic.csv?rlkey=ly36vi9v66sno80eir6rohlwn&dl=1", na = "(null)") |> # some values are coded as "(null)" in the df; rewrite them as NA
#   janitor::clean_names() 
df_nypd = read_csv("../NYPD_Complaint_Data_Historic.csv") |> 
  janitor::clean_names()

#Reading in the dataset that map each precinct to a neighborhood, for further SES analysis
prec_neighbor = read_csv("data/nyc_prec_neighborhood.csv")

#Merging `prec_neighboor` with NYPD dataset, conducting more data cleaning. The resulting `nypd_ses_df` only contains variables of interest for SES analysis. `nypd_ses_df` now contains neighboorhood information, as well as other necessary data from the original master data 
nypd_ses_df = df_nypd |> 
  select(cmplnt_num, cmplnt_fr_dt, addr_pct_cd, crm_atpt_cptd_cd, law_cat_cd, susp_age_group, susp_race, susp_sex, vic_age_group, vic_race, vic_sex, ofns_desc, pd_desc) |> 
  rename(precinct = addr_pct_cd) |> 
  mutate(cmplnt_fr_dt = as.Date(cmplnt_fr_dt, format = "%m/%d/%Y"),
         year = format(cmplnt_fr_dt, "%Y")) |> 
  left_join(prec_neighbor, by = "precinct")

#Reading in the SES data for demographics information for each precinct/neighboorhood, and only keeping the relevant SES indicators. `neighbor_ses` df includes information on population, income, umemployment, poverty, education, and racial factors for every neighborhood. 
neighbor_ses = readxl::read_excel("data/neighorhood_indicators.xlsx", sheet = "Data") |> 
  janitor::clean_names() |> 
  filter(region_type == "Sub-Borough Area") |> 
  rename(neighborhood = region_name) |> 
  select(neighborhood, year, pop_num, hh_inc_med_adj, pop16_unemp_pct, pop_edu_collp_pct, pop_edu_nohs_pct, pop_pov_pct, pop_race_asian_pct, pop_race_black_pct, pop_race_hisp_pct, pop_race_white_pct, pop_foreign_pct) |> 
  filter(year %in% c(2017, 2018, 2019, 2021)) #2020 and 2022 data missing

neighbor_rent = readxl::read_excel("data/neighorhood_indicators.xlsx", sheet = "Data") |> 
  janitor::clean_names() |> 
  filter(region_type == "Sub-Borough Area") |> 
  rename(neighborhood = region_name) |> 
  filter(year == "2017-2021") |> 
  select(neighborhood, gross_rent_0_1beds, gross_rent_2_3beds)

# Now, calculate the crime rate for every precinct. Crime rate is calculated by dividing the number of crimes committed per neighborhood over the total population of that neighborhood, then times 100,000. Then merge the crime rate data with other relevant socioeconomic data. We focus only on the 2021 data. `ses_df` contains crime rate and other SES data per neighborhood, by year 

ses_df = nypd_ses_df |> 
  group_by(year, borough, neighborhood) |> 
  summarise(crime_num = n())

ses_df = ses_df |> merge(neighbor_ses, by = c("year", "neighborhood")) |> 
  mutate(crime_rate = (crime_num/pop_num) * 100,000) |> 
  left_join(neighbor_rent, by = "neighborhood")
```

## Linear Regression

### Visualization of potential association

Visualizing potential association between poverty, income, and educational level with crime rate: 
```{r}
eda_plot1 = ses_df |> ggplot(aes(x = pop_pov_pct, y = crime_rate)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "% below poverty threshold vs. Crime Rate")

eda_plot2 = ses_df |> ggplot(aes(x = hh_inc_med_adj, y = crime_rate)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Median Household Income vs. Crime Rate")

eda_plot3 = ses_df |> ggplot(aes(x = pop16_unemp_pct, y = crime_rate)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "% umployed vs. Crime Rate")

eda_plot4 = ses_df |> ggplot(aes(x = pop_edu_nohs_pct, y = crime_rate)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "% without High School Education vs. Crime Rate")

gridExtra::grid.arrange(
  eda_plot1, eda_plot2, eda_plot3, eda_plot4, ncol = 2)
```

It is observed that there is a stronger positive association between percentage of poverty population with crime rate; a weaker positive association between unemployment and lack of education with crime rate, and a negative association between median household income and crime rate. 

### MLR model 

Let's first visualize the distribution of crime rate:
```{r}
ses_df |> ggplot(aes(x = crime_rate)) +
  geom_histogram(alpha = 0.8, color = "white") + 
  labs(
    x = "Crime Rate" ,
    y = "Count",
    title = "Distribution of Crime Rate by Neighborhoods")
```

Since the data is slightly right-skewed, we can use the Box-Cox Transformation to transform the data for linear fit
```{r}
initial_fit = lm(crime_rate ~ pop_pov_pct + hh_inc_med_adj + pop16_unemp_pct + pop_edu_nohs_pct, data = ses_df)

MASS::boxcox(initial_fit)
```

From the Box-Cox Transformation, log-transformation seems to be the best option. We will perform a log transformation on crime rate and use this as the predictor.
```{r}
ses_df = ses_df |> mutate(
  log_crime_rate = log(crime_rate)
)

log_fit = lm(log_crime_rate ~ pop_pov_pct + hh_inc_med_adj + pop16_unemp_pct + pop_edu_nohs_pct, data = ses_df)

broom::tidy(log_fit)
```

### MLR model diagnostics

From the result of the MLR, it is observed that all 4 predictors are significant in terms of association with the crime rate outcome. We can run a regression diagnostic by plotting residuals for each borough:
```{r}
ses_df |> 
  modelr::add_residuals(log_fit) |> 
  ggplot(aes(x = borough, y = resid)) + geom_violin() + 
  labs(title = "MLR residuals for each borough", 
       xlab = "Borough", 
       ylab = "Residuals")
```

We can further inspect our model by plotting residuals against fitted values
```{r}
ses_df |> 
  modelr::add_residuals(log_fit) |> 
  modelr::add_predictions(log_fit) |> 
  ggplot(aes(x = pred, y = resid, color = borough)) + geom_point() +
  labs(title = "MLR residuals against fitted value", 
       xlab = "Residuals", 
       ylab = "Fitted Values")
```

From the two diagnostics plots above, we can see that the residuals are roughly centered around zero, and are relatively normal. This would indicate that our model is a relatively good fit. 

## Lasso Regression

For the above MLR, we've only inspected four potential indicators that can predict crime rate. There are other social and demographics indicator that we are interested in, such as race and immigration status. Since Lasso regression encourages simpler model by only keeping the most influential factors, we also applied Lasso Regression to fine tune the prediction model.

```{r}
# Finding the best lambda
cv_model <- cv.glmnet(data.matrix(ses_df[, c('pop_pov_pct', 'hh_inc_med_adj', 'pop16_unemp_pct', 'pop_edu_nohs_pct', 'pop_race_asian_pct', 'pop_race_black_pct', 'pop_race_hisp_pct', 'pop_race_white_pct', 'pop_foreign_pct')]), ses_df$crime_rate, alpha = 1)

best_lambda <- cv_model$lambda.min

# Conducting the Lasso fit 
lasso_fit = glmnet(data.matrix(ses_df[, c('pop_pov_pct', 'hh_inc_med_adj', 'pop16_unemp_pct', 'pop_edu_nohs_pct', 'pop_race_asian_pct', 'pop_race_black_pct', 'pop_race_hisp_pct', 'pop_race_white_pct', 'pop_foreign_pct')]), ses_df$crime_rate, alpha = 1, lambda = best_lambda)

coef(lasso_fit) 
```

TO DO: add diagnostics and compare with MLR 