---
title: "Suspect and Victim Analysis"
author: "Miao Fu"
date: "2023-11-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message=FALSE,
  warning=FALSE)
library(tidyverse)
library(tidyr)
```

```{r,include=FALSE}
df_nypd = read_csv("https://www.dropbox.com/scl/fi/kf2zk4t1onxzm2vo3lpkq/NYPD_Complaint_Data_Historic.csv?rlkey=ly36vi9v66sno80eir6rohlwn&dl=1", na = "(null)") |> # some values are coded as "(null)" in the df; rewrite them as NA
  janitor::clean_names() 

demo=df_nypd|>
  select(boro_nm,susp_age_group,susp_race,susp_sex,vic_age_group,vic_race,vic_sex)|>
  mutate_all(~na_if(.,"UNKNOWN"))|>
  na.omit()
```

## Demographic Distribution

We would like to learn more about the demographic information breakdown in suspects and victims. The analysis is done through three aspects: race, age, and sex. We divided the exploration into different boroughs as it is a representative way of looking at NYC. 

### race distribution across boroughs

```{r, echo=FALSE}
susp_race=demo|>
  select(boro_nm,susp_race)|>
  group_by(boro_nm,susp_race)|>
  summarise(count=n())|>
  group_by(boro_nm)|>
  mutate(total=sum(count),
         percent=(count/total)*100)|>
  select(-count,-total)

colnames(susp_race)=c("boro_nm","race","percent_susp")

vic_race=demo|>
  select(boro_nm,vic_race)|>
  group_by(boro_nm,vic_race)|>
  summarise(count=n())|>
  group_by(boro_nm)|>
  mutate(total=sum(count),
         percent=(count/total)*100)|>
  select(-count,-total)

colnames(vic_race)=c("boro_nm","race","percent_vic")

combined_race=vic_race|>
  full_join(susp_race,by=c("boro_nm","race"))|>
  pivot_longer(
    cols=starts_with("percent"),
    names_to="category",
    values_to="percent"
  )|>
  mutate(
    category=ifelse(category=="percent_susp","SUSPECT","VICTIM")
  )

combined_race|>
  ggplot(aes(x=boro_nm,y=percent,fill=race))+
  geom_bar(stat="identity")+
  facet_grid(~category)+
  labs(
    title="Race Distribution in Suspects and Victims",
    x="NYC boroughs",
    y="Percentage(%)"
  )+
  theme_minimal()+
  scale_fill_brewer(palette = "Dark2")+
  theme(axis.text.x=element_text(angle=45,hjust=1),legend.position="bottom")
  
```

This figure shows the percentage of different race for all NYC boroughs. Black, white, and white hispanic are the top three percent in both suspects and victims across boroughs. The distribution of races is similar between suspects and victims for each borough. 

### age distribution across boroughs 
```{r, echo=FALSE}
susp_age=demo|>
  select(boro_nm,susp_age_group)|>
  group_by(boro_nm,susp_age_group)|>
  summarise(count=n())|>
  filter(count>10)|>
  group_by(boro_nm)|>
  mutate(total=sum(count),
         percent=(count/total)*100)|>
  select(-count,-total)

colnames(susp_age)=c("boro_nm","age","percent_susp")

vic_age=demo|>
  select(boro_nm,vic_age_group)|>
  group_by(boro_nm,vic_age_group)|>
  summarise(count=n())|>
  filter(count>10)|>
  group_by(boro_nm)|>
  mutate(total=sum(count),
         percent=(count/total)*100)|>
  select(-count,-total)

colnames(vic_age)=c("boro_nm","age","percent_vic")

combined_age=vic_age|>
  full_join(susp_age)|>
  pivot_longer(
    cols=starts_with("percent"),
    names_to="category",
    values_to="percent"
  )|>
  mutate(
    category=ifelse(category=="percent_susp","SUSPECT","VICTIM")
  )

combined_age|>
  ggplot(aes(x=boro_nm,y=percent,fill=age))+
  geom_bar(stat="identity")+
  facet_grid(~category)+
  labs(
    title="Age Distribution in Suspects and Victims",
    x="NYC boroughs",
    y="Percentage(%)"
  )+
  theme_minimal()+
  scale_fill_brewer(palette = "Dark2")+
  theme(axis.text.x=element_text(angle=45,hjust=1),legend.position="bottom")
```

This figure shows the percentage of age group for suspect and victim in all NYC boroughs. Majority of suspects and victims come from the age group of 25-44. The distribution of age is similar between suspects and victims for each borough as well as across boroughs. 

### sex distribution across boroughs 
```{r, echo=FALSE}
susp_sex=demo|>
  select(boro_nm,susp_sex)|>
  group_by(boro_nm,susp_sex)|>
  summarise(count=n())|>
  filter(susp_sex!="U")|>
  group_by(boro_nm)|>
  mutate(total=sum(count),
         percent=(count/total)*100)|>
  select(-count,-total)

colnames(susp_sex)=c("boro_nm","sex","percent_susp")

vic_sex=demo|>
  select(boro_nm,vic_sex)|>
  group_by(boro_nm,vic_sex)|>
  summarise(count=n())|>
  filter(vic_sex%in%c("F","M"))|>
  group_by(boro_nm)|>
  mutate(total=sum(count),
         percent=(count/total)*100)|>
  select(-count,-total)

colnames(vic_sex)=c("boro_nm","sex","percent_vic")

combined_sex=vic_sex|>
  full_join(susp_sex)|>
  pivot_longer(
    cols=starts_with("percent"),
    names_to="category",
    values_to="percent"
  )|>
  mutate(
    category=ifelse(category=="percent_susp","SUSPECT","VICTIM")
  )

combined_sex|>
  ggplot(aes(x=boro_nm,y=percent,fill=sex))+
  geom_bar(stat="identity")+
  facet_grid(~category)+
  labs(
    title="Sex Distribution in Suspects and Victims",
    x="NYC boroughs",
    y="Percentage(%)"
  )+
  theme_minimal()+
  scale_fill_brewer(palette = "Dark2")+
  theme(axis.text.x=element_text(angle=45,hjust=1),legend.position="bottom")
```

This figure shows the percentage of sex in suspects and victims for all NYC boroughs. Around 75% of suspects are male and around 25% are female for all boroughs. Around 70% of victims are female and around 30% are male for all boroughs. 


## Project Report paragaph

### Data 
   A subset of the master dataset was created specifically for demographic distribution analysis by selecting the relevant variables. Rows with missing values were dropped for data cleaning purpose. The demographic dataset is further divided into suspects and victims datasets to calculate the percentage of race, age and sex. Columns with incorrect data values are further cleaned up. The demographic dataset has following variables:
   
   * `boro_nm`: NYC borough name\
   * `susp_age_group`: Age group the suspect belongs to (<18, 18-24, 25-44, 45-64, 65+)\
   * `susp_race`: Suspect race (AMERICAN INDIAN/ALASKAN NATIVE, ASIAN / PACIFIC ISLANDER, BLACK, BLACK HISPANIC, WHITE, WHITE HISPANIC)\
   * `susp_sex`: Suspect sex (Female, Male)\
   * `vic_age_group`: Age group the victim belongs to (<18, 18-24, 25-44, 45-64, 65+)\
   * `vic_race`: Victim race (AMERICAN INDIAN/ALASKAN NATIVE, ASIAN / PACIFIC ISLANDER, BLACK, BLACK HISPANIC, WHITE, WHITE HISPANIC)\
   * `vic_sex`: Victim sex (Female, Male)

### Exploratory analysis
  Three barplots were created to visualize the distribution of race, age, and sex in suspects and victims. Number of suspects and number of victims are separately counted in each race, age, and sex groups. Initially, count of suspects and victims in each race, age, and sex groups were plotted. However, the different total population of each borough made it difficult to directly visualize the distributions. Thus, percentage of suspects and victims in each race, age, and sex groups were calculated by dividing the count by total number of suspects or victims in each borough. 

### Discussion
  It is interesting to see the distribution of race are quite similar between suspects and victims. This suggests that the racial composition of individuals involved as suspects or victims is consistent across the dataset. The distribution of race are also similar between boroughs. This suggests that there is no significant variation in the racial demographics of incidents across different boroughs. The distribution of age group are also similar between suspects and victims, as well as across boroughs. This implies consistency in age demographics across dataset and NYC boroughs. In combined, we observe consistency in the race and age group distribution in suspects and victims in NYC crimes. The distributions are uniform throughout the city, which was not expected. The uniformity in demographic patterns may indicate systemic factors influencing incidents rather than borough-specific or role-specific(suspect vs victim) factors. \
  
  On the other hand, the distribution of sex in suspects and victims show significantly more male suspects than female suspects across all 7 NYC boroughs. This suggests a consistent gender disparity among individuals identified as suspects. Additionally, there are significantly more female victims than male victims across all 7 NYC boroughs. In combined, the data highlights a gender imbalance with a higher proportion of males being identified as suspects and a higher proportion of females being victims. The gender distribution is not surprising and confirms what we expected. The observed gender disparities may suggests potential sociological or systemic factors influencing crime reporting, law enforcement, or victimization. 
