---
title: "Exploratory Data Visualization"
output: 
  html_document:
    keep_md: yes
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message=FALSE,
  warning=FALSE)
library(tidyverse)
library(dplyr)
library(sf)
library(tmap)
library(plotly)
```

While examining the NYPD crime complaint data, we expect crime rate to be indicative by the socioeconomic status of a specific neighborhood. In this section, we calculated the crime rate for each neighborhood each year, and displayed the top neighborhoods with highest crime rate, along with what their crime composition looks like. We are also interested in whether crime rate can be predictive of rent in corresponding neighborhoods. 


```{r}
# Import and clead NYPD data
# df_nypd = read_csv("https://www.dropbox.com/scl/fi/kf2zk4t1onxzm2vo3lpkq/NYPD_Complaint_Data_Historic.csv?rlkey=ly36vi9v66sno80eir6rohlwn&dl=1", na = "(null)") |> # some values are coded as "(null)" in the df; rewrite them as NA
#   janitor::clean_names() 

## Siqing' note: I'm currently reading from my local file. Yuki, when you run this file please comment out the chunk below and use the dropbox. Thanks!!

df_nypd = read_csv("../NYPD_Complaint_Data_Historic.csv") |> 
  janitor::clean_names()

prec_neighbor = read_csv("data/nyc_prec_neighborhood.csv")

#Merging neighborhood information with NYPD data, keep only relevant variables
nypd_ses_df = df_nypd |> 
  select(cmplnt_num, cmplnt_fr_dt, addr_pct_cd, crm_atpt_cptd_cd, law_cat_cd, susp_age_group, susp_race, susp_sex, vic_age_group, vic_race, vic_sex, ofns_desc, pd_desc, latitude, longitude) |> 
  rename(precinct = addr_pct_cd) |> 
  mutate(cmplnt_fr_dt = as.Date(cmplnt_fr_dt, format = "%m/%d/%Y"),
         year = format(cmplnt_fr_dt, "%Y")) |> 
  left_join(prec_neighbor, by = "precinct")

#Import and clean socio-economic data. neighbor_SES has SES indicators (except rent) for each neighborhood 
neighbor_ses = readxl::read_excel("data/neighorhood_indicators.xlsx", sheet = "Data") |> 
  janitor::clean_names() |> 
  filter(region_type == "Sub-Borough Area") |> 
  rename(neighborhood = region_name) |> 
  select(neighborhood, year, pop_num, hh_inc_med_adj, pop16_unemp_pct, pop_edu_collp_pct, pop_edu_nohs_pct, pop_pov_pct, pop_race_asian_pct, pop_race_black_pct, pop_race_hisp_pct, pop_race_white_pct, pop_foreign_pct) |> 
  filter(year %in% c(2017, 2018, 2019, 2020, 2021, 2022))

#neighbor_rent has rent data for each neighborhood 
neighbor_rent = readxl::read_excel("data/neighorhood_indicators.xlsx", sheet = "Data") |> 
  janitor::clean_names() |> 
  filter(region_type == "Sub-Borough Area") |> 
  rename(neighborhood = region_name) |> 
  filter(year == "2017-2021") |> 
  select(neighborhood, gross_rent_0_1beds, gross_rent_2_3beds)

#ses_df has crime rate and SES indicators for each neighborhood
ses_df = nypd_ses_df |> 
  group_by(year, borough, neighborhood) |> 
  summarise(crime_num = n())

ses_df = ses_df |> merge(neighbor_ses, by = c("year", "neighborhood")) |> 
  mutate(crime_rate = (crime_num/pop_num) * 100,000) |> 
  left_join(neighbor_rent, by = "neighborhood")
```

## Heatmap
```{r}
# NYC neighborhoods borders
nyc = read_sf(here::here("data", "Police_Precincts.geojson")) |> 
  select(-shape_area, -shape_leng) |> 
  mutate(
    precinct = as.double(precinct)
  )

# prepare data set for mapping
# population data by precinct (2020 census, P1_001N: total population)
df_pop = read_csv(here::here("data", "nyc_precinct_2020pop.csv")) |> 
  rename(pop = P1_001N) |> 
  select(precinct, pop)

# select ses variables
ses_heat = neighbor_ses |> 
  filter(year == 2021) |> # visualize 2021 data
  mutate(
    pop16_unemp_pct = pop16_unemp_pct * 100, 
    pop_edu_collp_pct = pop_edu_collp_pct * 100, 
    pop_pov_pct = pop_pov_pct * 100
  ) |> 
  select(neighborhood, pop_num, hh_inc_med_adj, pop16_unemp_pct, pop_edu_collp_pct, pop_pov_pct)

# merge with nypd dataset and calculate relevant indicators
nypd_ses_heat = nypd_ses_df |>  
  filter(year == 2021) |> 
  left_join(ses_heat, by = "neighborhood") |> # combine with ses data
  st_as_sf(
    # which columns to use as coordinates
    coords = c("longitude", "latitude"), 
    # keep the coordinate columns
    remove = FALSE,
    # projection system
    crs = 4326
  ) |> 
  merge(df_pop, by = "precinct") |> 
  group_by(precinct) |> 
  mutate(
    "crime_rate" = (sum(ofns_desc == "FELONY ASSAULT") / pop) * 100000, # calculate crime rate per 100,000 using 2020 population census data by precinct
    "violation" = (sum(law_cat_cd == "VIOLATION") / n()) * 100, # the rate of violation among all complaints
    "misdemeanor" = (sum(law_cat_cd == "MISDEMEANOR") / n()) * 100, # the rate of misdemeanor among all complaints
    "felony" = (sum(law_cat_cd == "FELONY") / n()) * 100 # the rate of felony among all complaints
  ) |> 
  select(precinct, neighborhood, hh_inc_med_adj, pop16_unemp_pct, pop_edu_collp_pct, pop_pov_pct, geometry, crime_rate, felony, misdemeanor, violation)

# spatial join
df_spj = nypd_ses_heat |>
  # spatial join
  st_join(
    # only need these columns from nyc tibble
    nyc |> select(precinct, geometry),
    # join rows where there is some overlap between a dock and a precinct
    join = st_intersects,
    left = FALSE
  ) |> 
  select(-precinct.y) |> 
  rename(precinct = precinct.x)

count_by_neighborhood = df_spj |> 
  # remove geometry for fast counting
  st_drop_geometry() |> 
  distinct() |> 
  # join the counts into the nyc neighborhood object
  right_join(nyc, by=c("precinct" = "precinct")) |> 
  mutate(
    "pre_nei" = paste(precinct, "-", neighborhood, sep = " ") 
  ) |> 
  st_as_sf() |> 
  drop_na()  # drop central park due to the lack of data

# tmap
tmap_mode("view") # set tmap to the view mode

tm_shape(count_by_neighborhood) +
  tm_polygons(col = "crime_rate", id = "pre_nei", breaks=c(0, 50, 100, 300, 600, 1000, 1600), title = "Crime rate per 100,000")

tm_shape(count_by_neighborhood) +
  tm_polygons(col = "hh_inc_med_adj", id = "pre_nei", title = "Median household’s income (USD/yr)", palette = "Blues") +
  tm_bubbles("crime_rate", id = "pre_nei", col = "azure", alpha = .5)

tm_shape(count_by_neighborhood) +
  tm_polygons(col = "pop16_unemp_pct", id = "pre_nei", title = "Unemployment rate (%)") +
  tm_bubbles("crime_rate", id = "pre_nei", col = "azure", alpha = .5)

tm_shape(count_by_neighborhood) +
  tm_polygons(col = "pop_edu_collp_pct", id = "pre_nei", title = "College graduates (%)", palette = "Blues") +
  tm_bubbles("crime_rate", id = "pre_nei", col = "azure", alpha = .5)

tm_shape(count_by_neighborhood) +
  tm_polygons(col = "pop_pov_pct", id = "pre_nei", title = "Below poverty threshold (%)") +
  tm_bubbles("crime_rate", id = "pre_nei", col = "azure", alpha = .5)
```

## Crime type

### What are the top 5 neighborhoods with highest average crime rate?
```{r}
ses_df |> group_by(borough, neighborhood) |>
  summarise(avg_crime_rate = mean(crime_rate, na.rm = TRUE)) |> 
  arrange(desc(avg_crime_rate)) |> head(5) |> knitr::kable()
```

### What are the top 5 neighborhoods with highest average crime rate?

```{r}
ses_df |> group_by(borough, neighborhood) |>
  summarise(avg_crime_rate = mean(crime_rate, na.rm = TRUE)) |> 
  arrange(desc(avg_crime_rate)) |> tail(5) |> knitr::kable()
```

Since CUMC is located in Manhattan, let's also visualize the crime rate in Manhattan:

```{r}
ses_df |> group_by(borough, neighborhood) |>
  filter(borough == "Manhattan") |> 
  summarise(avg_crime_rate = mean(crime_rate, na.rm = TRUE)) |> 
  arrange(desc(avg_crime_rate)) |> knitr::kable()
```

The top 5 neighborhoods of highest average crime rate are to our surprise, so we further looked into the common types of crimes in these neighorhoods. 

### What are the most common types of crime in these neighborhoods?

Chelsea/Clinton/Midtown:
```{r}
nypd_ses_df |> 
  filter(neighborhood =="Chelsea/Clinton/Midtown") |> 
  count(ofns_desc) |> 
  arrange(desc(n)) |> 
  head(5) |> 
  mutate(percentage = scales::percent(n / sum(n), scale = 100)) |> 
  select(ofns_desc, percentage) |> 
  rename(crime_type = ofns_desc) |> 
  knitr::kable()
```

University Heights/Fordham: 
```{r}
nypd_ses_df |> 
  filter(neighborhood == "University Heights/Fordham") |> 
  count(ofns_desc) |> 
  arrange(desc(n)) |> 
  head(5) |> 
  mutate(percentage = scales::percent(n / sum(n), scale = 100)) |> 
  select(ofns_desc, percentage) |> 
  rename(crime_type = ofns_desc) |>
  knitr::kable() 
```

Mott Haven/Hunts Point: 
```{r}
nypd_ses_df |> 
  filter(neighborhood == "Mott Haven/Hunts Point") |> 
  count(ofns_desc) |> 
  arrange(desc(n)) |> 
  head(5) |> 
  mutate(percentage = scales::percent(n / sum(n), scale = 100)) |> 
  select(ofns_desc, percentage) |> 
  rename(crime_type = ofns_desc) |>
  knitr::kable() 
```

East Harlem:
```{r}
nypd_ses_df |> 
  filter(neighborhood == "East Harlem") |> 
  count(ofns_desc) |> 
  arrange(desc(n)) |> 
  head(5) |> 
  mutate(percentage = scales::percent(n / sum(n), scale = 100)) |> 
  select(ofns_desc, percentage) |> 
  rename(crime_type = ofns_desc) |>
  knitr::kable() 
```

Lower East Side/Chinatown: 
```{r}
nypd_ses_df |> 
  filter(neighborhood == "Lower East Side/Chinatown") |> 
  count(ofns_desc) |> 
  arrange(desc(n)) |> 
  head(5) |> 
  mutate(percentage = scales::percent(n / sum(n), scale = 100)) |> 
  select(ofns_desc, percentage) |>
  rename(crime_type = ofns_desc) |>
  knitr::kable() 
```

The high crime rate in Chelsea/Clinton/Midtown can be explained by the significant contribution of petite and grand larceny, which is as expected. 

Since CUMC is located in Washington Heights, let's take a closer look at the top 10 crime types at Washington Heights: 

```{r}
nypd_ses_df |> 
  filter(neighborhood == "Washington Heights/Inwood") |> 
  count(ofns_desc) |> 
  arrange(desc(n)) |> 
  head(10) |> 
  mutate(percentage = n / sum(n) * 100) |>
  ggplot(aes(x = reorder(ofns_desc, -percentage), y = percentage)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) + 
   labs(title = "Top 10 types of crime in Washington Heights/Inwood",
       x = "Percentage of all crimes",
       y = "Type of crimes") 
```


## Line plot

### Is there an association between rent and crime rate?

```{r}
ses_df |> group_by(neighborhood) |> 
  summarise(avg_rent_0_1beds = mean(gross_rent_0_1beds), 
            avg_rent_2_3beds = mean(gross_rent_2_3beds)
            , avg_crime_rate = mean(crime_rate, na.rm = TRUE)) |> 
  ggplot(aes(x = avg_crime_rate)) +
  geom_point(aes(y = avg_rent_0_1beds, color = "Average Rent 0-1 Beds")) +
  geom_point(aes(y = avg_rent_2_3beds, color = "Average Rent 2-3 Beds")) +
  geom_smooth(aes(y = avg_rent_0_1beds), method = "lm", se = TRUE, color = "red") +
  geom_smooth(aes(y = avg_rent_2_3beds), method = "lm", se = TRUE, color = "blue") +
  scale_color_manual(values = c("Average Rent 0-1 Beds" = "red", "Average Rent 2-3 Beds" = "blue")) +
  labs(title = "Point Plot of average rent against average crime rate",
       x = "Crime rate",
       y = "Average rent ($)") 
```

There appears to be no obvious association between rent and crime rate.  

## Discussion



